---
title: "Metabolites Visualizations"
author: "Marc"
date: "15/4/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Libraries

```{r include=FALSE}
rm(list=ls()) 
set.seed(1234)
library(ggplot2)
library(ggfortify)
library(plotly)
library(cowplot)
library(caret)
library(readr)
library(tidyr)
library(dplyr)
library(purrr)
library(ggrepel)
library(patchwork)
library(ggpubr)
library(readxl)
library(reshape2)
library(tibble)
library(grid)
library(gridExtra)
library(lattice)
theme_set(theme_minimal())

plot_pca <- function(tmp1, tmp2, color_label, ranks = 6){
  
  tmp3 <- prcomp(tmp2, center = T, scale. = T, rank. = ranks)

  pdata <- 
    tmp3$x %>% 
    as_tibble() %>% 
    bind_cols(tmp1 %>% select(all_of(color_label)) %>% mutate_all(as.factor)) %>% 
    pivot_longer(cols = seq(1, ncol(.)-length(color_label), by=2)) %>% 
    select(x_val = value, x_name = name, everything()) %>% 
    pivot_longer(cols = starts_with("PC")) %>% 
    select(x_name, y_name = name,x_val, y_val = value, everything()) %>%
    mutate(name = paste(x_name, "vs", y_name),
           name = factor(name, levels = sort(name) %>% unique())) %>% 
    filter(as.numeric(gsub("PC", "", x_name)) + 1 == as.numeric(gsub("PC", "", y_name)))
  
  pca_plot <- pdata %>% 
    ggplot(aes(x=x_val, y=y_val))+
    geom_point(aes_string(fill = color_label), shape = 21, color = "white")+
    facet_grid(~name, scales = "free")+
    labs(x = "", y = "")+
    NULL
  return(pca_plot)
}
```

# Data loading

```{r}
WD <- getwd()
if(WD!="C:/Users/mysit/Documents/Bioinformatica/Semestre_4/workflowr/data"){
  setwd("../workflowr/data")
}

A_mdf_original<-readRDS("../data/mdf_original.rds")
B_mdf_4throot<-readRDS("../data/mdf_4throot.rds")
C_mdf_pca_outliers_removed<-readRDS("../data/mdf_pca_outliers_removed.rds")
D_mdf_extreme_values_removed<-readRDS("../data/mdf_extreme_values_removed.rds")
E_mdf_qc_removed<-readRDS("../data/mdf_qc_removed.rds")
F_mdf_sumtoone<-readRDS("../data/mdf_sumtoone.rds")
G_mdf_robustnorm<-readRDS("../data/mdf_robustnorm.rds")
H_mdf_meansnorm<-readRDS("../data/mdf_meansnorm.rds")
I_mdf_combat<-readRDS("../data/mdf_combat.rds")
J_mdf_wave_0<-readRDS("../data/mdf_wave_0.rds")
K_mdf_Rcpm<-readRDS("../data/mdf_Rcpm.rds")
L_mdf_MSBox<-readRDS("../data/mdf_MSBox.rds")


wave_0<-readRDS("../data/mdf_wave_0.rds")
wave_01<-readRDS("../data/mdf_wave_01.rds")
wave_02<-readRDS("../data/mdf_wave_02.rds")
wave_03<-readRDS("../data/mdf_wave_03.rds")
wave_04<-readRDS("../data/mdf_wave_04.rds")
wave_05<-readRDS("../data/mdf_wave_05.rds")
wave_06<-readRDS("../data/mdf_wave_06.rds")
wave_07<-readRDS("../data/mdf_wave_07.rds")
wave_08<-readRDS("../data/mdf_wave_08.rds")
wave_09<-readRDS("../data/mdf_wave_09.rds")
wave_1<-readRDS("../data/mdf_wave_1.rds")

sampleorder <- read_excel("../data/sample_order_traceage_wp2_sample_overview.xlsx") %>% as.data.frame()
sampleorder<-sampleorder %>% select(Sample, `Injection order positive`) 
names(sampleorder)[1]<-"sample"
mypalette<-c("#1296db","#c5dfed","#98e356","#8fb56d","#c4dead","#f77c1e","#ffc191","#FDBF6F")
```

# Identifying metabolites with high variance on the "before normalization" data
```{r}
#Finding a metabolite with high variance
variances <- apply(X=D_mdf_extreme_values_removed[7:(ncol(D_mdf_extreme_values_removed))], MARGIN=2, FUN=var)
sorted <- sort(variances, decreasing=TRUE, index.return=TRUE)$ix[1:20] # replace 2 with 100 ...
names(D_mdf_extreme_values_removed[, sorted])
```

# Plotting how the normalization methods change the batch effect

## Original data
```{r fig.asp = 0.65, fig.width = 6}
original <- merge(A_mdf_original,sampleorder,by=c("sample"), all = F) %>% relocate(`Injection order positive`, .before = sample)

poriginal<-original %>% ggplot(aes(`Injection order positive`,M91T29)) + geom_point(aes(colour = factor(batch))) + scale_color_manual(values = mypalette,breaks=c('1', 'b', '2','2b','b2','3','b3')) +labs(color = "Batch") + theme(axis.title.x = element_blank(),
          axis.title.y = element_blank())


original %>% ggplot(aes(`Injection order positive`,M91T29)) + geom_point(aes(colour = factor(batch))) + scale_color_manual(values = mypalette,breaks=c('1', 'b', '2','2b','b2','3','b3')) +labs(color = "Batch") + ggtitle("M91T29 intensity against injection order and batch on original unprocessed data") + theme(plot.title = element_text(face = "italic"))+  xlab("Injection Order") + ylab("Intensity")  +  xlab("Injection Order") + ylab("Intensity")
```

## After removing extreme values and applying 4th root transformation

```{r fig.asp = 0.65, fig.width = 6}
extremes <- merge(D_mdf_extreme_values_removed,sampleorder,by=c("sample"), all = F) %>% relocate(`Injection order positive`, .before = sample)

pextremes<-extremes %>% ggplot(aes(`Injection order positive`,M91T29)) + geom_point(aes(colour = factor(batch))) + scale_color_manual(values = mypalette,breaks=c('1', 'b', '2','2b','b2','3','b3')) +labs(color = "Batch") + theme(axis.title.x = element_blank(),
          axis.title.y = element_blank())


extremes %>% ggplot(aes(`Injection order positive`,M91T29)) + geom_point(aes(colour = factor(batch))) + scale_color_manual(values = mypalette,breaks=c('1', 'b', '2','2b','b2','3','b3')) +labs(color = "Batch") + ggtitle("Original data",subtitle = "M91T29 intensity by injection order and Batchs") + theme(plot.title = element_text(face = "italic"))+  xlab("Injection Order") + ylab("Intensity")+  xlab("Injection Order") + ylab("Intensity")
```

## After sum to 1 normalization

```{r fig.asp = 0.65, fig.width = 6}
sumto1 <- merge(F_mdf_sumtoone,sampleorder,by=c("sample"), all = F) %>% relocate(`Injection order positive`, .before = sample)

psumto1<-sumto1 %>% ggplot(aes(`Injection order positive`,M91T29)) + geom_point(aes(colour = factor(batch))) + scale_color_manual(values = mypalette,breaks=c('1', 'b', '2','2b','b2','3','b3')) +labs(color = "Batch") + theme(axis.title.x = element_blank(),
          axis.title.y = element_blank())

sumto1 %>% ggplot(aes(`Injection order positive`,M91T29)) + geom_point(aes(colour = factor(batch))) + scale_color_manual(values = mypalette,breaks=c('1', 'b', '2','2b','b2','3','b3')) +labs(color = "Batch") + ggtitle("Row normalisation",subtitle = "M91T29 intensity by injection order and Batchs") + theme(plot.title = element_text(face = "italic"))+  xlab("Injection Order") + ylab("Intensity")
```

## After "Rowbust" normalization

```{r fig.asp = 0.65, fig.width = 6}
robustnorm <- merge(G_mdf_robustnorm,sampleorder,by=c("sample"), all = F) %>% relocate(`Injection order positive`, .before = sample)

probust<-robustnorm %>% ggplot(aes(`Injection order positive`,M91T29)) + geom_point(aes(colour = factor(batch))) + scale_color_manual(values = mypalette,breaks=c('1', 'b', '2','2b','b2','3','b3')) +labs(color = "Batch") + theme(axis.title.x = element_blank(),
          axis.title.y = element_blank())

robustnorm %>% ggplot(aes(`Injection order positive`,M91T29)) + geom_point(aes(colour = factor(batch))) + scale_color_manual(values = mypalette,breaks=c('1', 'b', '2','2b','b2','3','b3')) +labs(color = "Batch") + ggtitle("Quantile normalisation",subtitle = "M91T29 intensity by injection order and Batchs") + theme(plot.title = element_text(face = "italic"))+  xlab("Injection Order") + ylab("Intensity")
```

## After mean centering normalization

```{r fig.asp = 0.65, fig.width = 6}
means <- merge(H_mdf_meansnorm,sampleorder,by=c("sample"), all = F) %>% relocate(`Injection order positive`, .before = sample)

pmeans<-means %>% ggplot(aes(`Injection order positive`,M91T29)) + geom_point(aes(colour = factor(batch))) + scale_color_manual(values = mypalette,breaks=c('1', 'b', '2','2b','b2','3','b3')) +labs(color = "Batch")+ theme(axis.title.x = element_blank(),
          axis.title.y = element_blank())


means %>% ggplot(aes(`Injection order positive`,M91T29)) + geom_point(aes(colour = factor(batch))) + scale_color_manual(values = mypalette,breaks=c('1', 'b', '2','2b','b2','3','b3')) +labs(color = "Batch") + ggtitle("Mean normalisation",subtitle = "M91T29 intensity by injection order and Batchs") + theme(plot.title = element_text(face = "italic"))+  xlab("Injection Order") + ylab("Intensity")
```

## After ComBat normalization

```{r fig.asp = 0.65, fig.width = 6}
combat <- merge(I_mdf_combat,sampleorder,by=c("sample"), all = F) %>% relocate(`Injection order positive`, .before = sample)

pcombat<-combat %>% ggplot(aes(`Injection order positive`,M91T29)) + geom_point(aes(colour = factor(batch))) + scale_color_manual(values = mypalette,breaks=c('1', 'b', '2','2b','b2','3','b3')) +labs(color = "Batch")+ theme(axis.title.x = element_blank(),
          axis.title.y = element_blank())
combat %>% ggplot(aes(`Injection order positive`,M91T29)) + geom_point(aes(colour = factor(batch))) + scale_color_manual(values = mypalette,breaks=c('1', 'b', '2','2b','b2','3','b3')) +labs(color = "Batch") + ggtitle("ComBat",subtitle = "M91T29 intensity by injection order and Batchs") + theme(plot.title = element_text(face = "italic"))+  xlab("Injection Order") + ylab("Intensity")
```

## After WaveICA using 0

```{r fig.asp = 0.65, fig.width = 6}
wave <- merge(J_mdf_wave_0,sampleorder,by=c("sample"), all = F) %>% relocate(`Injection order positive`, .before = sample)

pwave<-wave %>% ggplot(aes(`Injection order positive`,M91T29)) + geom_point(aes(colour = factor(batch))) + scale_color_manual(values = mypalette,breaks=c('1', 'b', '2','2b','b2','3','b3')) +labs(color = "Batch")+ theme(axis.title.x = element_blank(),
          axis.title.y = element_blank())
wave %>% ggplot(aes(`Injection order positive`,M91T29)) + geom_point(aes(colour = factor(batch))) + scale_color_manual(values = mypalette,breaks=c('1', 'b', '2','2b','b2','3','b3')) +labs(color = "Batch") + ggtitle("WaveICA2.0 cutoff 0",subtitle = "M91T29 intensity by injection order and Batchs") + theme(plot.title = element_text(face = "italic"))+  xlab("Injection Order") + ylab("Intensity")
```

## After PQN Rcpm

```{r fig.asp = 0.65, fig.width = 6}
rcpm <- merge(K_mdf_Rcpm,sampleorder,by=c("sample"), all = F) %>% relocate(`Injection order positive`, .before = sample)

prcpm<-rcpm %>% ggplot(aes(`Injection order positive`,M91T29)) + geom_point(aes(colour = factor(batch))) + scale_color_manual(values = mypalette,breaks=c('1', 'b', '2','2b','b2','3','b3')) +labs(color = "Batch")+ theme(axis.title.x = element_blank(),
          axis.title.y = element_blank())

rcpm %>% ggplot(aes(`Injection order positive`,M91T29)) + geom_point(aes(colour = factor(batch))) + scale_color_manual(values = mypalette,breaks=c('1', 'b', '2','2b','b2','3','b3')) +labs(color = "Batch") + ggtitle("PQN Rcpm",subtitle = "M91T29 intensity by injection order and Batchs") + theme(plot.title = element_text(face = "italic"))+  xlab("Injection Order") + ylab("Intensity")
```

## After PQN MSBox

```{r fig.asp = 0.65, fig.width = 6}
msbox <- merge(L_mdf_MSBox,sampleorder,by=c("sample"), all = F) %>% relocate(`Injection order positive`, .before = sample)

pmsbox<-msbox %>% ggplot(aes(`Injection order positive`,M91T29)) + geom_point(aes(colour = factor(batch))) + scale_color_manual(values = mypalette,breaks=c('1', 'b', '2','2b','b2','3','b3')) +labs(color = "Batch")+ theme(axis.title.x = element_blank(),
          axis.title.y = element_blank())
msbox %>% ggplot(aes(`Injection order positive`,M91T29)) + geom_point(aes(colour = factor(batch))) + scale_color_manual(values = mypalette,breaks=c('1', 'b', '2','2b','b2','3','b3')) +labs(color = "Batch") + ggtitle("PQN MSBox",subtitle = "M91T29 intensity by injection order and Batchs") + theme(plot.title = element_text(face = "italic"))+  xlab("Injection Order") + ylab("Intensity")
```


```{r fig.asp = 0.7, fig.width = 12}
y.grob <- textGrob("Metabolite (M91T29) Intensity", 
                   gp=gpar(), rot=90)

x.grob <- textGrob("Injection Order", 
                   gp=gpar())
arrangement<-ggarrange(pextremes, psumto1, probust,pmeans,pcombat, pwave,prcpm,pmsbox,
                    labels = c("Original data processed", "Row normalisation", "Quotient normalisation","Means normalisation","ComBat","WaveICA2.0 cut.0","PQN Rcpm","PQN MSBox"),font.label = list(size = 8, color = "black"),
                    ncol = 3, nrow = 3,common.legend = TRUE, legend = "right")



arrangement<-grid.arrange(arrangeGrob(arrangement, left = y.grob, bottom = x.grob))
annotate_figure(arrangement, top = text_grob("Injection order and intensity of a specific feature (M91T29)", face = "italic", size = 15))

rm(arrangement)
```

# WaveICA cutoff

```{r}
iwave0 <- merge(wave_0,sampleorder,by=c("sample"), all = F) %>% relocate(`Injection order positive`, .before = sample)

pwave0<-iwave0 %>% ggplot(aes(`Injection order positive`,M91T29)) + geom_point(aes(colour = factor(batch))) + scale_color_manual(values = mypalette,breaks=c('1', 'b', '2','2b','b2','3','b3')) +labs(color = "Batch")+ theme(axis.title.x = element_blank(),
          axis.title.y = element_blank())


iwave0 %>% ggplot(aes(`Injection order positive`,M91T29)) + geom_point(aes(colour = factor(batch))) + scale_color_manual(values = mypalette,breaks=c('1', 'b', '2','2b','b2','3','b3')) +labs(color = "Batch") + ggtitle("WaveICA2.0 cutoff of 0",subtitle = "M91T29 intensity by injection order and Batchs") + theme(plot.title = element_text(face = "italic"))+  xlab("Injection Order") + ylab("Intensity")

###
iwave01 <- merge(wave_01,sampleorder,by=c("sample"), all = F) %>% relocate(`Injection order positive`, .before = sample)

pwave01<-iwave01 %>% ggplot(aes(`Injection order positive`,M91T29)) + geom_point(aes(colour = factor(batch))) + scale_color_manual(values = mypalette,breaks=c('1', 'b', '2','2b','b2','3','b3')) +labs(color = "Batch")+ theme(axis.title.x = element_blank(),
          axis.title.y = element_blank())
###
iwave02 <- merge(wave_02,sampleorder,by=c("sample"), all = F) %>% relocate(`Injection order positive`, .before = sample)

pwave02<-iwave02 %>% ggplot(aes(`Injection order positive`,M91T29)) + geom_point(aes(colour = factor(batch))) + scale_color_manual(values = mypalette,breaks=c('1', 'b', '2','2b','b2','3','b3')) +labs(color = "Batch")+ theme(axis.title.x = element_blank(),
          axis.title.y = element_blank())
###
iwave03 <- merge(wave_03,sampleorder,by=c("sample"), all = F) %>% relocate(`Injection order positive`, .before = sample)

pwave03<-iwave03 %>% ggplot(aes(`Injection order positive`,M91T29)) + geom_point(aes(colour = factor(batch))) + scale_color_manual(values = mypalette,breaks=c('1', 'b', '2','2b','b2','3','b3')) +labs(color = "Batch")+ theme(axis.title.x = element_blank(),
          axis.title.y = element_blank())
###
iwave04 <- merge(wave_04,sampleorder,by=c("sample"), all = F) %>% relocate(`Injection order positive`, .before = sample)

pwave04<-iwave04 %>% ggplot(aes(`Injection order positive`,M91T29)) + geom_point(aes(colour = factor(batch))) + scale_color_manual(values = mypalette,breaks=c('1', 'b', '2','2b','b2','3','b3')) +labs(color = "Batch")+ theme(axis.title.x = element_blank(),
          axis.title.y = element_blank())
###
iwave05 <- merge(wave_05,sampleorder,by=c("sample"), all = F) %>% relocate(`Injection order positive`, .before = sample)

pwave05<-iwave05 %>% ggplot(aes(`Injection order positive`,M91T29)) + geom_point(aes(colour = factor(batch))) + scale_color_manual(values = mypalette,breaks=c('1', 'b', '2','2b','b2','3','b3')) +labs(color = "Batch")+ theme(axis.title.x = element_blank(),
          axis.title.y = element_blank())

iwave05 %>% ggplot(aes(`Injection order positive`,M91T29)) + geom_point(aes(colour = factor(batch))) + scale_color_manual(values = mypalette,breaks=c('1', 'b', '2','2b','b2','3','b3')) +labs(color = "Batch") + ggtitle("WaveICA2.0 cutoff of 0.5",subtitle = "M91T29 intensity by injection order and Batchs") + theme(plot.title = element_text(face = "italic"))+  xlab("Injection Order") + ylab("Intensity")
###
iwave06 <- merge(wave_06,sampleorder,by=c("sample"), all = F) %>% relocate(`Injection order positive`, .before = sample)

pwave06<-iwave06 %>% ggplot(aes(`Injection order positive`,M91T29)) + geom_point(aes(colour = factor(batch))) + scale_color_manual(values = mypalette,breaks=c('1', 'b', '2','2b','b2','3','b3')) +labs(color = "Batch")+ theme(axis.title.x = element_blank(),
          axis.title.y = element_blank())
###
iwave07 <- merge(wave_07,sampleorder,by=c("sample"), all = F) %>% relocate(`Injection order positive`, .before = sample)

pwave07<-iwave07 %>% ggplot(aes(`Injection order positive`,M91T29)) + geom_point(aes(colour = factor(batch))) + scale_color_manual(values = mypalette,breaks=c('1', 'b', '2','2b','b2','3','b3')) +labs(color = "Batch")+ theme(axis.title.x = element_blank(),
          axis.title.y = element_blank())
###
iwave08 <- merge(wave_08,sampleorder,by=c("sample"), all = F) %>% relocate(`Injection order positive`, .before = sample)

pwave08<-iwave08 %>% ggplot(aes(`Injection order positive`,M91T29)) + geom_point(aes(colour = factor(batch))) + scale_color_manual(values = mypalette,breaks=c('1', 'b', '2','2b','b2','3','b3')) +labs(color = "Batch")+ theme(axis.title.x = element_blank(),
          axis.title.y = element_blank())
###
iwave09 <- merge(wave_09,sampleorder,by=c("sample"), all = F) %>% relocate(`Injection order positive`, .before = sample)

pwave09<-iwave09 %>% ggplot(aes(`Injection order positive`,M91T29)) + geom_point(aes(colour = factor(batch))) + scale_color_manual(values = mypalette,breaks=c('1', 'b', '2','2b','b2','3','b3')) +labs(color = "Batch")+ theme(axis.title.x = element_blank(),
          axis.title.y = element_blank())
###
iwave1 <- merge(wave_1,sampleorder,by=c("sample"), all = F) %>% relocate(`Injection order positive`, .before = sample)

pwave1<-iwave1 %>% ggplot(aes(`Injection order positive`,M91T29)) + geom_point(aes(colour = factor(batch))) + scale_color_manual(values = mypalette,breaks=c('1', 'b', '2','2b','b2','3','b3')) +labs(color = "Batch")+ theme(axis.title.x = element_blank(),
          axis.title.y = element_blank())

iwave1 %>% ggplot(aes(`Injection order positive`,M91T29)) + geom_point(aes(colour = factor(batch))) + scale_color_manual(values = mypalette,breaks=c('1', 'b', '2','2b','b2','3','b3')) +labs(color = "Batch") + ggtitle("WaveICA2.0 cutoff of 1",subtitle = "M91T29 intensity by injection order and Batchs") + theme(plot.title = element_text(face = "italic"))+  xlab("Injection Order") + ylab("Intensity")
```

```{r fig.asp = 0.7, fig.width = 12}
y.grob <- textGrob("Metabolite (M91T29) Intensity", 
                   gp=gpar(), rot=90)

x.grob <- textGrob("Injection Order", 
                   gp=gpar())
arrangement<-ggarrange(pwave1, pwave09, pwave08,pwave07, pwave06,pwave05,pwave04,pwave03,pwave02,pwave01,pwave0,
                    labels = c("Cut = 1", "Cut = 0.9", "Cut = 0.8","Cut = 0.7","Cut = 0.6","Cut = 0.5","Cut = 0.4","Cut = 0.3","Cut = 0.2","Cut = 0.1","Cut = 0"),font.label = list(size = 8, color = "black"),
                    ncol = 4, nrow = 3,common.legend = TRUE, legend = "right")



arrangement<-grid.arrange(arrangeGrob(arrangement, left = y.grob, bottom = x.grob))
annotate_figure(arrangement, top = text_grob("Injection order and intensity of a specific feature (M91T29) by WaveICA2.0 cutoff parameter", face = "italic", size = 15))

rm(arrangement)
```


```{r fig.asp = 0.7, fig.width = 12}
y.grob <- textGrob("Metabolite (M91T29) Intensity", 
                   gp=gpar(), rot=90)

x.grob <- textGrob("Injection Order", 
                   gp=gpar())
arrangement<-ggarrange(pwave1,pwave05,pwave0,
                    labels = c("Cut = 1","Cut = 0.5","Cut = 0"),font.label = list(size = 8, color = "black"),
                    ncol = 2, nrow = 2,common.legend = TRUE, align = "hv",legend = "right")



arrangement<-grid.arrange(arrangeGrob(arrangement, left = y.grob, bottom = x.grob))
annotate_figure(arrangement, top = text_grob("Injection order and intensity of a specific feature (M91T29) by WaveICA2.0 cutoff parameter", face = "italic", size = 15))

rm(arrangement)
```
