---
title: "RMSE Visualizations RF Ranger"
author: "Marc"
date: "16/4/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Libraries

```{r include=FALSE}
rm(list=ls()) 
set.seed(1234)
library(caret)
library(cowplot)
library(dplyr)
library(forcats)
library(ggfortify)
library(ggplot2)
library(ggpubr)
library(ggrepel)
#
library(grid)
#
library(gridExtra)
library(lattice)
library(patchwork)
library(plotly)
library(purrr)
library(readr)
library(readxl)
library(reshape2)
library(tibble)
library(tidyr)
theme_set(theme_minimal())
```

# Model loading
```{r}
WD <- getwd()
if(WD!="C:/Users/mysit/Documents/Bioinformatica/Semestre_4/workflowr/data"){
  setwd("../workflowr/data")
}

load("models_ranger.RData")
```

# Errors and plot from the original data until extreme values removed

## Original Data

```{r}
results <- data.frame(TrainRMSE = as.numeric(),TrainRsquared = as.numeric(),TrainMAE = as.numeric(),method = as.character(),Name = as.character(),Details = as.character())


results[1,] = data.frame(getTrainPerf(A_Input_data), Name = paste0("A_Input_Data"), Details = "Original Raw Data")

error<-A_Input_data$pred %>%
  filter(mtry == A_Input_data$bestTune$mtry & splitrule == A_Input_data$bestTune$splitrule & min.node.size == A_Input_data$bestTune$min.node.size) %>% #subset 
  separate(Resample, c("fold", "rep"), "\\.") %>% 
  group_by(rep) %>% #group by replicate
  summarise(RMSE = RMSE(obs, pred)) %>% 
  as.data.frame() %>% 
  mutate(Model = "A_Input_data", Details = "Original Raw Data")
```

## 4th root

```{r}
results[2,] = data.frame(getTrainPerf(B_4th_root), Name = paste0("B_4th_root"), Details = "4th root transformed")

error1<-B_4th_root$pred %>%
  filter(mtry == B_4th_root$bestTune$mtry & splitrule == B_4th_root$bestTune$splitrule & min.node.size == B_4th_root$bestTune$min.node.size) %>% #subset 
  separate(Resample, c("fold", "rep"), "\\.") %>% 
  group_by(rep) %>% #group by replicate
  summarise(RMSE = RMSE(obs, pred)) %>% 
  as.data.frame() %>% 
  mutate(Model = "B_4th_root", Details = "4th root transformed")

error<-rbind(error,error1)
```

## PCA outliers removed

```{r}
results[3,] = data.frame(getTrainPerf(C_PCA_outliers_removed), Name = paste0("C_PCA_outliers_removed"), Details = "PCA outliers removed")

error1<-C_PCA_outliers_removed$pred %>%
  filter(mtry == C_PCA_outliers_removed$bestTune$mtry & splitrule == C_PCA_outliers_removed$bestTune$splitrule & min.node.size == C_PCA_outliers_removed$bestTune$min.node.size) %>% #subset 
  separate(Resample, c("fold", "rep"), "\\.") %>% 
  group_by(rep) %>% #group by replicate
  summarise(RMSE = RMSE(obs, pred)) %>% 
  as.data.frame() %>% 
  mutate(Model = "C_PCA_outliers_removed", Details = "PCA outliers removed")

error<-rbind(error,error1)
```

## Extreme values removed

```{r}
results[4,] = data.frame(getTrainPerf(D_extreme_vals_removed), Name = paste0("D_extreme_vals_removed"), Details = "Extreme values removed")

error1<-D_extreme_vals_removed$pred %>%
  filter(mtry == D_extreme_vals_removed$bestTune$mtry & splitrule == D_extreme_vals_removed$bestTune$splitrule & min.node.size == D_extreme_vals_removed$bestTune$min.node.size) %>% #subset 
  separate(Resample, c("fold", "rep"), "\\.") %>% 
  group_by(rep) %>% #group by replicate
  summarise(RMSE = RMSE(obs, pred)) %>% 
  as.data.frame() %>% 
  mutate(Model = "D_extreme_vals_removed", Details = "Extreme values removed")

error<-rbind(error,error1)
```

## QC cleanup

```{r}
results[5,] = data.frame(getTrainPerf(E_QC_cleanup), Name = paste0("E_QC_cleanup"), Details = "QC Cleanup")

error1<-E_QC_cleanup$pred %>%
  filter(mtry == E_QC_cleanup$bestTune$mtry & splitrule == E_QC_cleanup$bestTune$splitrule & min.node.size == E_QC_cleanup$bestTune$min.node.size) %>% #subset 
  separate(Resample, c("fold", "rep"), "\\.") %>% 
  group_by(rep) %>% #group by replicate
  summarise(RMSE = RMSE(obs, pred)) %>% 
  as.data.frame() %>% 
  mutate(Model = "E_QC_cleanup", Details = "QC Cleanup")

error<-rbind(error,error1)
```


## Making the cool plot
```{r fig.asp = 0.9, fig.width = 7}
#write_rds(error,"error_init_rf.RDS")
error<-error %>% filter(Details!="QC Cleanup")
results<-results %>% filter(Details!="QC Cleanup")
error.p<-error %>% ggplot(aes(x=fct_inorder(Details),y=RMSE, color=Model))+
  geom_boxplot(outlier.colour = "red", alpha = 0.3)+
  geom_boxplot(outlier.colour="red", outlier.shape=1,
                outlier.size=8)+
  geom_jitter(shape=16, position=position_jitter(0))+
  theme(legend.position="top")+
  scale_color_brewer(palette = "Set2")+
  theme(axis.title.x = element_blank())

table.p <- ggtexttable(results, rows = NULL, 
                        theme = ttheme("light", base_size = 7))
ggarrange(error.p, table.p, 
          ncol = 1, nrow = 2)

```

```{r fig.asp = 0.65, fig.width = 6}
table.p
```


# Errors and plot from the normalization methods

## Extreme values again

```{r}
results <- data.frame(TrainRMSE = as.numeric(),TrainRsquared = as.numeric(),TrainMAE = as.numeric(),method = as.character(),Name = as.character(),Details = as.character())

results[1,] = data.frame(getTrainPerf(D_extreme_vals_removed), Name = paste0("D_extreme_vals_removed"), Details = "Extreme values removed")

error<-D_extreme_vals_removed$pred %>%
  filter(mtry == D_extreme_vals_removed$bestTune$mtry & splitrule == D_extreme_vals_removed$bestTune$splitrule & min.node.size == D_extreme_vals_removed$bestTune$min.node.size) %>% #subset 
  separate(Resample, c("fold", "rep"), "\\.") %>% 
  group_by(rep) %>% #group by replicate
  summarise(RMSE = RMSE(obs, pred)) %>% 
  as.data.frame() %>% 
  mutate(Model = "D_extreme_vals_removed", Details = "Extreme values removed")
```

## Sum to 1

```{r}
results[2,] = data.frame(getTrainPerf(F_sum_to_1_norm), Name = paste0("F_sum_to_1_norm"), Details = "RowsSum to 1")

error1<-F_sum_to_1_norm$pred %>%
  filter(mtry == F_sum_to_1_norm$bestTune$mtry & splitrule == F_sum_to_1_norm$bestTune$splitrule & min.node.size == F_sum_to_1_norm$bestTune$min.node.size) %>% #subset 
  separate(Resample, c("fold", "rep"), "\\.") %>% 
  group_by(rep) %>% #group by replicate
  summarise(RMSE = RMSE(obs, pred)) %>% 
  as.data.frame() %>% 
  mutate(Model = "F_sum_to_1_norm", Details = "RowsSum to 1")

error<-rbind(error,error1)
```

## Robust

```{r}
results[3,] = data.frame(getTrainPerf(G_og_norm), Name = paste0("G_og_norm"), Details = "Robust")

error1<-G_og_norm$pred %>%
  filter(mtry == G_og_norm$bestTune$mtry & splitrule == G_og_norm$bestTune$splitrule & min.node.size == G_og_norm$bestTune$min.node.size) %>% #subset 
  separate(Resample, c("fold", "rep"), "\\.") %>% 
  group_by(rep) %>% #group by replicate
  summarise(RMSE = RMSE(obs, pred)) %>% 
  as.data.frame() %>% 
  mutate(Model = "G_og_norm", Details = "Robust")

error<-rbind(error,error1)
```

## Means

```{r}
results[4,] = data.frame(getTrainPerf(H_BN_means), Name = paste0("H_BN_means"), Details = "Mean Centering")

error1<-H_BN_means$pred %>%
  filter(mtry == H_BN_means$bestTune$mtry & splitrule == H_BN_means$bestTune$splitrule & min.node.size == H_BN_means$bestTune$min.node.size) %>% #subset 
  separate(Resample, c("fold", "rep"), "\\.") %>% 
  group_by(rep) %>% #group by replicate
  summarise(RMSE = RMSE(obs, pred)) %>% 
  as.data.frame() %>% 
  mutate(Model = "H_BN_means", Details = "Mean Centering")

error<-rbind(error,error1)
```

## ComBat

```{r}
results[5,] = data.frame(getTrainPerf(I_ComBat), Name = paste0("I_ComBat"), Details = "ComBat")

error1<-I_ComBat$pred %>%
  filter(mtry == I_ComBat$bestTune$mtry & splitrule == I_ComBat$bestTune$splitrule & min.node.size == I_ComBat$bestTune$min.node.size) %>% #subset 
  separate(Resample, c("fold", "rep"), "\\.") %>% 
  group_by(rep) %>% #group by replicate
  summarise(RMSE = RMSE(obs, pred)) %>% 
  as.data.frame() %>% 
  mutate(Model = "I_ComBat", Details = "ComBat")

error<-rbind(error,error1)
```

## WaveICA2.0 Cutoff of 0

```{r}
results[6,] = data.frame(getTrainPerf(wave_0), Name = paste0("wave_0"), Details = "WaveICA2.0 Cutoff 0")

error1<-wave_0$pred %>%
  filter(mtry == wave_0$bestTune$mtry & splitrule == wave_0$bestTune$splitrule & min.node.size == wave_0$bestTune$min.node.size) %>% #subset 
  separate(Resample, c("fold", "rep"), "\\.") %>% 
  group_by(rep) %>% #group by replicate
  summarise(RMSE = RMSE(obs, pred)) %>% 
  as.data.frame() %>% 
  mutate(Model = "wave_0", Details = "WaveICA2.0 Cutoff 0")

error<-rbind(error,error1)
```

## PQN Rcpm

```{r}
results[7,] = data.frame(getTrainPerf(J_PQN_Rcpm), Name = paste0("J_PQN_Rcpm"), Details = "PQN Rcpm")

error1<-J_PQN_Rcpm$pred %>%
  filter(mtry == J_PQN_Rcpm$bestTune$mtry & splitrule == J_PQN_Rcpm$bestTune$splitrule & min.node.size == J_PQN_Rcpm$bestTune$min.node.size) %>% #subset 
  separate(Resample, c("fold", "rep"), "\\.") %>% 
  group_by(rep) %>% #group by replicate
  summarise(RMSE = RMSE(obs, pred)) %>% 
  as.data.frame() %>% 
  mutate(Model = "J_PQN_Rcpm", Details = "PQN Rcpm")

error<-rbind(error,error1)
```


## PQN msBox

```{r}
results[8,] = data.frame(getTrainPerf(K_PQN_msBox), Name = paste0("K_PQN_msBox"), Details = "PQN msBox")

error1<-K_PQN_msBox$pred %>%
  filter(mtry == K_PQN_msBox$bestTune$mtry & splitrule == K_PQN_msBox$bestTune$splitrule & min.node.size == K_PQN_msBox$bestTune$min.node.size) %>% #subset 
  separate(Resample, c("fold", "rep"), "\\.") %>% 
  group_by(rep) %>% #group by replicate
  summarise(RMSE = RMSE(obs, pred)) %>% 
  as.data.frame() %>% 
  mutate(Model = "K_PQN_msBox", Details = "PQN msBox")

error<-rbind(error,error1)
```

## Making the cool plot
```{r fig.asp = 0.7, fig.width = 9}
#saveRDS(error,"error_models_RF.RDS")
error.p<-error %>% ggplot(aes(x=fct_inorder(Details),y=RMSE, color=Model))+
  geom_boxplot(outlier.colour = "red", alpha = 0.3)+
  geom_boxplot(outlier.colour="red", outlier.shape=1,
                outlier.size=8)+
  geom_jitter(shape=16, position=position_jitter(0))+
  theme(legend.position="top")+
  scale_color_brewer(palette = "Set2")+
  theme(axis.title.x = element_blank())

table.p <- ggtexttable(results, rows = NULL, 
                        theme = ttheme("light", base_size = 7))
ggarrange(error.p, table.p, 
          ncol = 1, nrow = 2,
          heights = c(0.8, 0.5, 0.3))

ggtexttable(results, rows = NULL, 
                        theme = ttheme("light"))
results_rf<-results
```


# Errors and plot from the WaveICA2.0 Cutoff parameters

## WaveICA2.0 Cutoff of 0

```{r}
results <- data.frame(TrainRMSE = as.numeric(),TrainRsquared = as.numeric(),TrainMAE = as.numeric(),method = as.character(),Name = as.character(),Details = as.character())

results[1,] = data.frame(getTrainPerf(wave_0), Name = paste0("wave_0"), Details = "WaveICA2.0 Cutoff 0")

error<-wave_0$pred %>%
  filter(mtry == wave_0$bestTune$mtry & splitrule == wave_0$bestTune$splitrule & min.node.size == wave_0$bestTune$min.node.size) %>% #subset 
  separate(Resample, c("fold", "rep"), "\\.") %>% 
  group_by(rep) %>% #group by replicate
  summarise(RMSE = RMSE(obs, pred)) %>% 
  as.data.frame() %>% 
  mutate(Model = "wave_0", Details = "WaveICA2.0 Cutoff 0")
```

## WaveICA2.0 Cutoff of 0.1

```{r}
results[2,] = data.frame(getTrainPerf(wave_0.1), Name = paste0("wave_0.1"), Details = "WaveICA2.0 Cutoff 0.1")

error1<-wave_0.1$pred %>%
  filter(mtry == wave_0.1$bestTune$mtry & splitrule == wave_0.1$bestTune$splitrule & min.node.size == wave_0.1$bestTune$min.node.size) %>% #subset 
  separate(Resample, c("fold", "rep"), "\\.") %>% 
  group_by(rep) %>% #group by replicate
  summarise(RMSE = RMSE(obs, pred)) %>% 
  as.data.frame() %>% 
  mutate(Model = "wave_0.1", Details = "WaveICA2.0 Cutoff 0.1")

error<-rbind(error,error1)
```

## WaveICA2.0 Cutoff of 0.2

```{r}
results[3,] = data.frame(getTrainPerf(wave_0.2), Name = paste0("wave_0.2"), Details = "WaveICA2.0 Cutoff 0.2")

error1<-wave_0.2$pred %>%
  filter(mtry == wave_0.2$bestTune$mtry & splitrule == wave_0.2$bestTune$splitrule & min.node.size == wave_0.2$bestTune$min.node.size) %>% #subset 
  separate(Resample, c("fold", "rep"), "\\.") %>% 
  group_by(rep) %>% #group by replicate
  summarise(RMSE = RMSE(obs, pred)) %>% 
  as.data.frame() %>% 
  mutate(Model = "wave_0.2", Details = "WaveICA2.0 Cutoff 0.2")

error<-rbind(error,error1)
```

## WaveICA2.0 Cutoff of 0.3

```{r}
results[4,] = data.frame(getTrainPerf(wave_0.3), Name = paste0("wave_0.3"), Details = "WaveICA2.0 Cutoff 0.3")

error1<-wave_0.3$pred %>%
  filter(mtry == wave_0.3$bestTune$mtry & splitrule == wave_0.3$bestTune$splitrule & min.node.size == wave_0.3$bestTune$min.node.size) %>% #subset 
  separate(Resample, c("fold", "rep"), "\\.") %>% 
  group_by(rep) %>% #group by replicate
  summarise(RMSE = RMSE(obs, pred)) %>% 
  as.data.frame() %>% 
  mutate(Model = "wave_0.3", Details = "WaveICA2.0 Cutoff 0.3")

error<-rbind(error,error1)
```

## WaveICA2.0 Cutoff of 0.4

```{r}
results[5,] = data.frame(getTrainPerf(wave_0.4), Name = paste0("wave_0.4"), Details = "WaveICA2.0 Cutoff 0.4")

error1<-wave_0.4$pred %>%
  filter(mtry == wave_0.4$bestTune$mtry & splitrule == wave_0.4$bestTune$splitrule & min.node.size == wave_0.4$bestTune$min.node.size) %>% #subset 
  separate(Resample, c("fold", "rep"), "\\.") %>% 
  group_by(rep) %>% #group by replicate
  summarise(RMSE = RMSE(obs, pred)) %>% 
  as.data.frame() %>% 
  mutate(Model = "wave_0.4", Details = "WaveICA2.0 Cutoff 0.4")

error<-rbind(error,error1)
```

## WaveICA2.0 Cutoff of 0.5

```{r}
results[6,] = data.frame(getTrainPerf(wave_0.5), Name = paste0("wave_0.5"), Details = "WaveICA2.0 Cutoff 0.5")

error1<-wave_0.5$pred %>%
  filter(mtry == wave_0.5$bestTune$mtry & splitrule == wave_0.5$bestTune$splitrule & min.node.size == wave_0.5$bestTune$min.node.size) %>% #subset 
  separate(Resample, c("fold", "rep"), "\\.") %>% 
  group_by(rep) %>% #group by replicate
  summarise(RMSE = RMSE(obs, pred)) %>% 
  as.data.frame() %>% 
  mutate(Model = "wave_0.5", Details = "WaveICA2.0 Cutoff 0.5")

error<-rbind(error,error1)
```


## WaveICA2.0 Cutoff of 0.6

```{r}
results[7,] = data.frame(getTrainPerf(wave_0.6), Name = paste0("wave_0.6"), Details = "WaveICA2.0 Cutoff 0.6")

error1<-wave_0.6$pred %>%
  filter(mtry == wave_0.6$bestTune$mtry & splitrule == wave_0.6$bestTune$splitrule & min.node.size == wave_0.6$bestTune$min.node.size) %>% #subset 
  separate(Resample, c("fold", "rep"), "\\.") %>% 
  group_by(rep) %>% #group by replicate
  summarise(RMSE = RMSE(obs, pred)) %>% 
  as.data.frame() %>% 
  mutate(Model = "wave_0.6", Details = "WaveICA2.0 Cutoff 0.6")

error<-rbind(error,error1)
```

## WaveICA2.0 Cutoff of 0.7

```{r}
results[8,] = data.frame(getTrainPerf(wave_0.7), Name = paste0("wave_0.7"), Details = "WaveICA2.0 Cutoff 0.7")

error1<-wave_0.7$pred %>%
  filter(mtry == wave_0.7$bestTune$mtry & splitrule == wave_0.7$bestTune$splitrule & min.node.size == wave_0.7$bestTune$min.node.size) %>% #subset 
  separate(Resample, c("fold", "rep"), "\\.") %>% 
  group_by(rep) %>% #group by replicate
  summarise(RMSE = RMSE(obs, pred)) %>% 
  as.data.frame() %>% 
  mutate(Model = "wave_0.7", Details = "WaveICA2.0 Cutoff 0.7")

error<-rbind(error,error1)
```

## WaveICA2.0 Cutoff of 0.8

```{r}
results[9,] = data.frame(getTrainPerf(wave_0.8), Name = paste0("wave_0.8"), Details = "WaveICA2.0 Cutoff 0.8")

error1<-wave_0.8$pred %>%
  filter(mtry == wave_0.8$bestTune$mtry & splitrule == wave_0.8$bestTune$splitrule & min.node.size == wave_0.8$bestTune$min.node.size) %>% #subset 
  separate(Resample, c("fold", "rep"), "\\.") %>% 
  group_by(rep) %>% #group by replicate
  summarise(RMSE = RMSE(obs, pred)) %>% 
  as.data.frame() %>% 
  mutate(Model = "wave_0.8", Details = "WaveICA2.0 Cutoff 0.8")

error<-rbind(error,error1)
```

## WaveICA2.0 Cutoff of 0.9

```{r}
results[10,] = data.frame(getTrainPerf(wave_0.9), Name = paste0("wave_0.9"), Details = "WaveICA2.0 Cutoff 0.9")

error1<-wave_0.9$pred %>%
  filter(mtry == wave_0.9$bestTune$mtry & splitrule == wave_0.9$bestTune$splitrule & min.node.size == wave_0.9$bestTune$min.node.size) %>% #subset 
  separate(Resample, c("fold", "rep"), "\\.") %>% 
  group_by(rep) %>% #group by replicate
  summarise(RMSE = RMSE(obs, pred)) %>% 
  as.data.frame() %>% 
  mutate(Model = "wave_0.9", Details = "WaveICA2.0 Cutoff 0.9")

error<-rbind(error,error1)
```

## WaveICA2.0 Cutoff of 1

```{r}
results[11,] = data.frame(getTrainPerf(wave_1), Name = paste0("wave_1"), Details = "WaveICA2.0 Cutoff 1")

error1<-wave_1$pred %>%
  filter(mtry == wave_1$bestTune$mtry & splitrule == wave_1$bestTune$splitrule & min.node.size == wave_1$bestTune$min.node.size) %>% #subset 
  separate(Resample, c("fold", "rep"), "\\.") %>% 
  group_by(rep) %>% #group by replicate
  summarise(RMSE = RMSE(obs, pred)) %>% 
  as.data.frame() %>% 
  mutate(Model = "wave_1", Details = "WaveICA2.0 Cutoff 1")

error<-rbind(error,error1)
```

## Making the cool plots
```{r fig.asp = 1, fig.width = 14}
error.p<-error %>% ggplot(aes(x=fct_inorder(Details),y=RMSE, color=Model))+
  geom_boxplot(outlier.colour = "red", alpha = 0.3)+
  geom_boxplot(outlier.colour="red", outlier.shape=1,
                outlier.size=8)+
  geom_jitter(shape=16, position=position_jitter(0))+
  theme(legend.position="top")+
  theme(axis.title.x = element_blank())

table.p <- ggtexttable(results, rows = NULL, 
                        theme = ttheme("light", base_size = 7))
ggarrange(error.p, table.p, 
          ncol = 1, nrow = 2,
          heights = c(1, 0.5, 0.3))
```

```{r fig.asp = 1, fig.width = 14}
error.p<-error %>% slice(26:nrow(error)) %>% ggplot(aes(x=fct_inorder(Details),y=RMSE, color=Model))+
  geom_boxplot(outlier.colour = "red", alpha = 0.3)+
  geom_boxplot(outlier.colour="red", outlier.shape=1,
                outlier.size=8)+
  geom_jitter(shape=16, position=position_jitter(0))+
  theme(legend.position="top")+
  theme(axis.title.x = element_blank())

resultstemp<-results %>% slice(2:nrow(results))
table.p <- ggtexttable(resultstemp, rows = NULL, 
                        theme = ttheme("light", base_size = 7))
rm(resultstemp)
ggarrange(error.p, table.p, 
          ncol = 1, nrow = 2,
          heights = c(1, 0.5, 0.3))
```

```{r}
#saveRDS(error, "error_rf_wave.RDS")
```

# Making plots combining both models

## Init
```{r}
error_init_glmnet<-readRDS("error_init_glmnet.RDS")
error_init_rf<-readRDS("error_init_rf.RDS")
results_glmnet<-readRDS("results_glmnet.RDS")

error_init_rf$Model <- paste0(error_init_rf$Model,"_RF")
error_init_rf$type <- "Ranger"
error_init_glmnet$Model <- paste0(error_init_glmnet$Model,"_GLMNET")
error_init_glmnet$type <- "GLMNET"

error_init <- rbind(error_init_rf,error_init_glmnet)
error_init<-error_init[order(error_init$Model),]
```

```{r fig.asp = 0.65, fig.width = 6}
error_init<-error_init %>% filter(Details!="QC Cleanup")

error_init %>% ggplot(aes(x=fct_inorder(Details),y=RMSE, color=type))+
  stat_boxplot(geom ='errorbar', width = 0.6)+
  geom_jitter(shape=16,position = position_jitterdodge(jitter.width = 0,
  jitter.height = 0,dodge.width = 0.6),alpha=0.4,size=2)+
  geom_boxplot(width = 0.6,outlier.colour = "red", alpha = 0.6, outlier.shape = 18, outlier.size = 3)+
  geom_boxplot(width = 0.6,outlier.shape=1,
                outlier.size=6,alpha=.5,outlier.colour="red")+
  theme(axis.title.x = element_blank()) +
  scale_x_discrete(guide = guide_axis(n.dodge=2))+
  scale_color_brewer(palette = "Set2")+ 
  labs(color='Model')+
  theme(plot.title = element_text(face = "italic"))+
  theme(legend.position="right")+
  ggtitle("Data preparation", subtitle = "CARET model performance, Boxplot")
```

## Models

```{r}
error_models_RF<-readRDS("error_models_RF.RDS")
error_models_GLMNET<-readRDS("error_models_GLMNET.RDS")

error_models_RF$Model <- paste0(error_models_RF$Model,"_RF")
error_models_RF$type <- "Ranger"
error_models_GLMNET$Model <- paste0(error_models_GLMNET$Model,"_GLMNET")
error_models_GLMNET$type <- "GLMNET"

error_models <- rbind(error_models_RF,error_models_GLMNET)
error_models<-error_models[order(error_models$Model),]
```

### Original data

```{r fig.asp = 0.65, fig.width = 6}
full_results<-rbind(results_glmnet,results_rf)
full_results<-full_results[order(full_results$Name),]


error.p<-error_models %>% slice(1:50) %>% ggplot(aes(x=fct_inorder(Model),y=RMSE, color=type))+
  geom_boxplot(outlier.colour = "red", alpha = 0.3)+
  geom_boxplot(outlier.colour="red", outlier.shape=1,
                outlier.size=8,alpha=0.3)+
  geom_jitter(shape=16,position = position_jitter(height = 0, width = 0))+
  theme(legend.position="right")+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  scale_x_discrete(guide = guide_axis(n.dodge=3))+ 
  ggtitle("Model performance on original data")+
  theme(plot.title = element_text(face = "italic"))+ labs(color = "Model")+scale_color_brewer(palette="Set2")

table.p <- ggtexttable(full_results[1:2,], rows = NULL, 
                        theme = ttheme("light", base_size = 7))

ggarrange(error.p, table.p, 
          ncol = 1, nrow = 2,
          heights = c(1, 0.5, 1))

```

### Whole plot

```{r fig.asp = 0.6, fig.width = 8}
error_models %>% ggplot(aes(x=fct_inorder(Details),y=RMSE, color=type))+
  stat_boxplot(geom ='errorbar', width = 0.6)+
  geom_jitter(shape=16,position = position_jitterdodge(jitter.width = 0,
  jitter.height = 0,dodge.width = 0.6),alpha=0.4,size=2)+
  geom_boxplot(width = 0.6,outlier.colour = "red", alpha = 0.6, outlier.shape = 18, outlier.size = 3)+
  geom_boxplot(width = 0.6,outlier.shape=1,
                outlier.size=6,alpha=.5,outlier.colour="red")+
  theme(legend.position="top")+
  theme(axis.title.x = element_blank()) +
  scale_x_discrete(guide = guide_axis(n.dodge=2))+
  scale_color_brewer(palette = "Set2")+ 
  labs(color='Model')+
  theme(plot.title = element_text(face = "italic"))+
  theme(legend.position="right")+
  ggtitle("All batch correction methods", subtitle = "CARET model performance, Boxplot")
```

```{r fig.asp = 0.65, fig.width = 6}
ggtexttable(full_results, rows = NULL, 
                        theme = ttheme("light", base_size = 7))
```


## Wave

```{r}
error_rf_wave <- readRDS("error_rf_wave.RDS")
error_glmnet_wave <- readRDS("error_glmnet_wave.RDS")

error_rf_wave$Model <- paste0(error_rf_wave$Model,"_RF")
error_rf_wave$type <- "Ranger"
error_glmnet_wave$Model <- paste0(error_glmnet_wave$Model,"_GLMNET")
error_glmnet_wave$type <- "GLMNET"


error_wave <- rbind(error_rf_wave,error_glmnet_wave)
error_wave<-error_wave[order(error_wave$Model),]
```

```{r fig.asp = 0.6, fig.width = 8}
error_wave <- error_wave %>% 
        mutate(Details = sub(".*? ", "", error_wave$Details))
error_wave <- error_wave %>% 
        mutate(Details = sub(".*? ", "", error_wave$Details))

error_wave %>% ggplot(aes(x=fct_inorder(Details),y=RMSE, color=type))+
  stat_boxplot(geom ='errorbar', width = 0.6)+
  geom_jitter(shape=16,position = position_jitterdodge(jitter.width = 0,
  jitter.height = 0,dodge.width = 0.6),alpha=0.4,size=2)+
  geom_boxplot(width = 0.6,outlier.colour = "red", alpha = 0.6, outlier.shape = 18, outlier.size = 3)+
  geom_boxplot(width = 0.6,outlier.shape=1,
                outlier.size=6,alpha=.5,outlier.colour="red")+
  theme(legend.position="top")+
  theme(axis.title.x = element_blank()) +
  scale_x_discrete(guide = guide_axis(n.dodge=2))+
  scale_color_brewer(palette = "Set2")+ 
  labs(color='Model')+
  theme(plot.title = element_text(face = "italic"))+
  theme(legend.position="right")+
  ggtitle("WaveICA2.0 cutoff study", subtitle = "CARET model performance, Boxplot")

results_wave_glmnet<-read_rds("../data/results_wave_glmnet.RDS")

full_results_wave<-rbind(results_wave_glmnet,results)
full_results_wave<-full_results_wave[order(full_results_wave$Name),]

# resultstemp<-results %>% slice(2:nrow(results))
# table.p <- ggtexttable(resultstemp, rows = NULL, 
#                         theme = ttheme("light", base_size = 7))
# rm(resultstemp)
# ggarrange(error.p, table.p, 
#           ncol = 1, nrow = 2,
#           heights = c(1, 0.5, 0.3))
```

```{r fig.asp = 1, fig.width = 6}
ggtexttable(full_results_wave, rows = NULL, 
                        theme = ttheme("light", base_size = 7))
```
