---
title: "RMSE Visualizations"
author: "Marc"
date: "16/4/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Libraries

```{r include=FALSE}
rm(list=ls()) 
set.seed(1234)
library(ggplot2)
library(ggfortify)
library(plotly)
library(cowplot)
library(caret)
library(readr)
library(tidyr)
library(dplyr)
library(purrr)
library(ggrepel)
library(patchwork)
library(ggpubr)
library(readxl)
library(reshape2)
library(tibble)
library(grid)
library(gridExtra)
library(lattice)
library(forcats)
theme_set(theme_minimal())
```

# Model loading
```{r}
WD <- getwd()
if(WD!="C:/Users/mysit/Documents/Bioinformatica/Semestre_4/workflowr/data"){
  setwd("../workflowr/data")
}

load("models.RData")
```

# Erros and plot from the original data until extreme values removed

## Original Data

```{r}
results <- data.frame(TrainRMSE = as.numeric(),TrainRsquared = as.numeric(),TrainMAE = as.numeric(),method = as.character(),Name = as.character(),Details = as.character())


results[1,] = data.frame(getTrainPerf(A_Input_data), Name = paste0("A_Input_Data"), Details = "Original Raw Data")

error<-A_Input_data$pred %>%
  filter(alpha == A_Input_data$bestTune$alpha & lambda == A_Input_data$bestTune$lambda) %>% #subset 
  separate(Resample, c("fold", "rep"), "\\.") %>% 
  group_by(rep) %>% #group by replicate
  summarise(RMSE = RMSE(obs, pred)) %>% 
  as.data.frame() %>% 
  mutate(Model = "A_Input_data", Details = "Original Raw Data")
```

## 4th root

```{r}
results[2,] = data.frame(getTrainPerf(B_4th_root), Name = paste0("B_4th_root"), Details = "4th root transformed")

error1<-B_4th_root$pred %>%
  filter(alpha == B_4th_root$bestTune$alpha & lambda == B_4th_root$bestTune$lambda) %>% #subset 
  separate(Resample, c("fold", "rep"), "\\.") %>% 
  group_by(rep) %>% #group by replicate
  summarise(RMSE = RMSE(obs, pred)) %>% 
  as.data.frame() %>% 
  mutate(Model = "B_4th_root", Details = "4th root transformed")

error<-rbind(error,error1)
```

## PCA outliers removed

```{r}
results[3,] = data.frame(getTrainPerf(C_PCA_outliers_removed), Name = paste0("C_PCA_outliers_removed"), Details = "PCA outliers removed")

error1<-C_PCA_outliers_removed$pred %>%
  filter(alpha == C_PCA_outliers_removed$bestTune$alpha & lambda == C_PCA_outliers_removed$bestTune$lambda) %>% #subset 
  separate(Resample, c("fold", "rep"), "\\.") %>% 
  group_by(rep) %>% #group by replicate
  summarise(RMSE = RMSE(obs, pred)) %>% 
  as.data.frame() %>% 
  mutate(Model = "C_PCA_outliers_removed", Details = "PCA outliers removed")

error<-rbind(error,error1)
```

## Extreme values removed

```{r}
results[4,] = data.frame(getTrainPerf(D_extreme_vals_removed), Name = paste0("D_extreme_vals_removed"), Details = "Extreme values removed")

error1<-D_extreme_vals_removed$pred %>%
  filter(alpha == D_extreme_vals_removed$bestTune$alpha & lambda == D_extreme_vals_removed$bestTune$lambda) %>% #subset 
  separate(Resample, c("fold", "rep"), "\\.") %>% 
  group_by(rep) %>% #group by replicate
  summarise(RMSE = RMSE(obs, pred)) %>% 
  as.data.frame() %>% 
  mutate(Model = "D_extreme_vals_removed", Details = "Extreme values removed")

error<-rbind(error,error1)
```

## QC cleanup

```{r}
results[5,] = data.frame(getTrainPerf(E_QC_cleanup), Name = paste0("E_QC_cleanup"), Details = "QC Cleanup")

error1<-E_QC_cleanup$pred %>%
  filter(alpha == E_QC_cleanup$bestTune$alpha & lambda == E_QC_cleanup$bestTune$lambda) %>% #subset 
  separate(Resample, c("fold", "rep"), "\\.") %>% 
  group_by(rep) %>% #group by replicate
  summarise(RMSE = RMSE(obs, pred)) %>% 
  as.data.frame() %>% 
  mutate(Model = "E_QC_cleanup", Details = "QC Cleanup")

error<-rbind(error,error1)
```


## Making the cool plot
```{r fig.asp = 0.9, fig.width = 7}
error.p<-error %>% ggplot(aes(x=fct_inorder(Details),y=RMSE, color=Model))+
  geom_boxplot(outlier.colour = "red", alpha = 0.3)+
  geom_boxplot(outlier.colour="red", outlier.shape=1,
                outlier.size=8)+
  geom_jitter(shape=16, position=position_jitter(0))+
  theme(legend.position="top")+
  scale_color_brewer(palette = "Set2")+
  theme(axis.title.x = element_blank())

table.p <- ggtexttable(results, rows = NULL, 
                        theme = ttheme("mOrange", base_size = 7))
ggarrange(error.p, table.p, 
          ncol = 1, nrow = 2)
```




# Errors and plot from the normalization methods

## Extreme values again

```{r}
results <- data.frame(TrainRMSE = as.numeric(),TrainRsquared = as.numeric(),TrainMAE = as.numeric(),method = as.character(),Name = as.character(),Details = as.character())

results[1,] = data.frame(getTrainPerf(D_extreme_vals_removed), Name = paste0("D_extreme_vals_removed"), Details = "Extreme values removed")

error<-D_extreme_vals_removed$pred %>%
  filter(alpha == D_extreme_vals_removed$bestTune$alpha & lambda == D_extreme_vals_removed$bestTune$lambda) %>% #subset 
  separate(Resample, c("fold", "rep"), "\\.") %>% 
  group_by(rep) %>% #group by replicate
  summarise(RMSE = RMSE(obs, pred)) %>% 
  as.data.frame() %>% 
  mutate(Model = "D_extreme_vals_removed", Details = "Extreme values removed")
```

## Sum to 1

```{r}
results[2,] = data.frame(getTrainPerf(F_sum_to_1_norm), Name = paste0("F_sum_to_1_norm"), Details = "RowsSum to 1")

error1<-F_sum_to_1_norm$pred %>%
  filter(alpha == F_sum_to_1_norm$bestTune$alpha & lambda == F_sum_to_1_norm$bestTune$lambda) %>% #subset 
  separate(Resample, c("fold", "rep"), "\\.") %>% 
  group_by(rep) %>% #group by replicate
  summarise(RMSE = RMSE(obs, pred)) %>% 
  as.data.frame() %>% 
  mutate(Model = "F_sum_to_1_norm", Details = "RowsSum to 1")

error<-rbind(error,error1)
```

## Robust

```{r}
results[3,] = data.frame(getTrainPerf(G_og_norm), Name = paste0("G_og_norm"), Details = "Robust")

error1<-G_og_norm$pred %>%
  filter(alpha == G_og_norm$bestTune$alpha & lambda == G_og_norm$bestTune$lambda) %>% #subset 
  separate(Resample, c("fold", "rep"), "\\.") %>% 
  group_by(rep) %>% #group by replicate
  summarise(RMSE = RMSE(obs, pred)) %>% 
  as.data.frame() %>% 
  mutate(Model = "G_og_norm", Details = "Robust")

error<-rbind(error,error1)
```

## Means

```{r}
results[4,] = data.frame(getTrainPerf(H_BN_means), Name = paste0("H_BN_means"), Details = "Mean Centering")

error1<-H_BN_means$pred %>%
  filter(alpha == H_BN_means$bestTune$alpha & lambda == H_BN_means$bestTune$lambda) %>% #subset 
  separate(Resample, c("fold", "rep"), "\\.") %>% 
  group_by(rep) %>% #group by replicate
  summarise(RMSE = RMSE(obs, pred)) %>% 
  as.data.frame() %>% 
  mutate(Model = "H_BN_means", Details = "Mean Centering")

error<-rbind(error,error1)
```

## ComBat

```{r}
results[5,] = data.frame(getTrainPerf(I_ComBat), Name = paste0("I_ComBat"), Details = "ComBat")

error1<-I_ComBat$pred %>%
  filter(alpha == I_ComBat$bestTune$alpha & lambda == I_ComBat$bestTune$lambda) %>% #subset 
  separate(Resample, c("fold", "rep"), "\\.") %>% 
  group_by(rep) %>% #group by replicate
  summarise(RMSE = RMSE(obs, pred)) %>% 
  as.data.frame() %>% 
  mutate(Model = "I_ComBat", Details = "ComBat")

error<-rbind(error,error1)
```

## WaveICA2.0 Cutoff of 0

```{r}
results[6,] = data.frame(getTrainPerf(wave_0), Name = paste0("wave_0"), Details = "WaveICA2.0 Cutoff 0")

error1<-wave_0$pred %>%
  filter(alpha == wave_0$bestTune$alpha & lambda == wave_0$bestTune$lambda) %>% #subset 
  separate(Resample, c("fold", "rep"), "\\.") %>% 
  group_by(rep) %>% #group by replicate
  summarise(RMSE = RMSE(obs, pred)) %>% 
  as.data.frame() %>% 
  mutate(Model = "wave_0", Details = "WaveICA2.0 Cutoff 0")

error<-rbind(error,error1)
```

## PQN Rcpm

```{r}
results[7,] = data.frame(getTrainPerf(J_PQN_Rcpm), Name = paste0("J_PQN_Rcpm"), Details = "PQN Rcpm")

error1<-J_PQN_Rcpm$pred %>%
  filter(alpha == J_PQN_Rcpm$bestTune$alpha & lambda == J_PQN_Rcpm$bestTune$lambda) %>% #subset 
  separate(Resample, c("fold", "rep"), "\\.") %>% 
  group_by(rep) %>% #group by replicate
  summarise(RMSE = RMSE(obs, pred)) %>% 
  as.data.frame() %>% 
  mutate(Model = "J_PQN_Rcpm", Details = "PQN Rcpm")

error<-rbind(error,error1)
```


## PQN msBox

```{r}
results[8,] = data.frame(getTrainPerf(K_PQN_msBox), Name = paste0("K_PQN_msBox"), Details = "PQN msBox")

error1<-K_PQN_msBox$pred %>%
  filter(alpha == K_PQN_msBox$bestTune$alpha & lambda == K_PQN_msBox$bestTune$lambda) %>% #subset 
  separate(Resample, c("fold", "rep"), "\\.") %>% 
  group_by(rep) %>% #group by replicate
  summarise(RMSE = RMSE(obs, pred)) %>% 
  as.data.frame() %>% 
  mutate(Model = "K_PQN_msBox", Details = "PQN msBox")

error<-rbind(error,error1)
```

## Making the cool plot
```{r fig.asp = 0.7, fig.width = 9}
error.p<-error %>% ggplot(aes(x=fct_inorder(Details),y=RMSE, color=Model))+
  geom_boxplot(outlier.colour = "red", alpha = 0.3)+
  geom_boxplot(outlier.colour="red", outlier.shape=1,
                outlier.size=8)+
  geom_jitter(shape=16, position=position_jitter(0))+
  theme(legend.position="top")+
  scale_color_brewer(palette = "Set2")+
  theme(axis.title.x = element_blank())

table.p <- ggtexttable(results, rows = NULL, 
                        theme = ttheme("mOrange", base_size = 7))
ggarrange(error.p, table.p, 
          ncol = 1, nrow = 2,
          heights = c(0.8, 0.5, 0.3))
```


# Errors and plot from the WaveICA2.0 Cutoff parameters

## WaveICA2.0 Cutoff of 0

```{r}
results <- data.frame(TrainRMSE = as.numeric(),TrainRsquared = as.numeric(),TrainMAE = as.numeric(),method = as.character(),Name = as.character(),Details = as.character())

results[1,] = data.frame(getTrainPerf(wave_0), Name = paste0("wave_0"), Details = "WaveICA2.0 Cutoff 0")

error<-wave_0$pred %>%
  filter(alpha == wave_0$bestTune$alpha & lambda == wave_0$bestTune$lambda) %>% #subset 
  separate(Resample, c("fold", "rep"), "\\.") %>% 
  group_by(rep) %>% #group by replicate
  summarise(RMSE = RMSE(obs, pred)) %>% 
  as.data.frame() %>% 
  mutate(Model = "wave_0", Details = "WaveICA2.0 Cutoff 0")
```

## WaveICA2.0 Cutoff of 0.1

```{r}
results[2,] = data.frame(getTrainPerf(wave_0.1), Name = paste0("wave_0.1"), Details = "WaveICA2.0 Cutoff 0.1")

error1<-wave_0.1$pred %>%
  filter(alpha == wave_0.1$bestTune$alpha & lambda == wave_0.1$bestTune$lambda) %>% #subset 
  separate(Resample, c("fold", "rep"), "\\.") %>% 
  group_by(rep) %>% #group by replicate
  summarise(RMSE = RMSE(obs, pred)) %>% 
  as.data.frame() %>% 
  mutate(Model = "wave_0.1", Details = "WaveICA2.0 Cutoff 0.1")

error<-rbind(error,error1)
```

## WaveICA2.0 Cutoff of 0.2

```{r}
results[3,] = data.frame(getTrainPerf(wave_0.2), Name = paste0("wave_0.2"), Details = "WaveICA2.0 Cutoff 0.2")

error1<-wave_0.2$pred %>%
  filter(alpha == wave_0.2$bestTune$alpha & lambda == wave_0.2$bestTune$lambda) %>% #subset 
  separate(Resample, c("fold", "rep"), "\\.") %>% 
  group_by(rep) %>% #group by replicate
  summarise(RMSE = RMSE(obs, pred)) %>% 
  as.data.frame() %>% 
  mutate(Model = "wave_0.2", Details = "WaveICA2.0 Cutoff 0.2")

error<-rbind(error,error1)
```

## WaveICA2.0 Cutoff of 0.3

```{r}
results[4,] = data.frame(getTrainPerf(wave_0.3), Name = paste0("wave_0.3"), Details = "WaveICA2.0 Cutoff 0.3")

error1<-wave_0.3$pred %>%
  filter(alpha == wave_0.3$bestTune$alpha & lambda == wave_0.3$bestTune$lambda) %>% #subset 
  separate(Resample, c("fold", "rep"), "\\.") %>% 
  group_by(rep) %>% #group by replicate
  summarise(RMSE = RMSE(obs, pred)) %>% 
  as.data.frame() %>% 
  mutate(Model = "wave_0.3", Details = "WaveICA2.0 Cutoff 0.3")

error<-rbind(error,error1)
```

## WaveICA2.0 Cutoff of 0.4

```{r}
results[5,] = data.frame(getTrainPerf(wave_0.4), Name = paste0("wave_0.4"), Details = "WaveICA2.0 Cutoff 0.4")

error1<-wave_0.4$pred %>%
  filter(alpha == wave_0.4$bestTune$alpha & lambda == wave_0.4$bestTune$lambda) %>% #subset 
  separate(Resample, c("fold", "rep"), "\\.") %>% 
  group_by(rep) %>% #group by replicate
  summarise(RMSE = RMSE(obs, pred)) %>% 
  as.data.frame() %>% 
  mutate(Model = "wave_0.4", Details = "WaveICA2.0 Cutoff 0.4")

error<-rbind(error,error1)
```

## WaveICA2.0 Cutoff of 0.5

```{r}
results[6,] = data.frame(getTrainPerf(wave_0.5), Name = paste0("wave_0.5"), Details = "WaveICA2.0 Cutoff 0.5")

error1<-wave_0.5$pred %>%
  filter(alpha == wave_0.5$bestTune$alpha & lambda == wave_0.5$bestTune$lambda) %>% #subset 
  separate(Resample, c("fold", "rep"), "\\.") %>% 
  group_by(rep) %>% #group by replicate
  summarise(RMSE = RMSE(obs, pred)) %>% 
  as.data.frame() %>% 
  mutate(Model = "wave_0.5", Details = "WaveICA2.0 Cutoff 0.5")

error<-rbind(error,error1)
```


## WaveICA2.0 Cutoff of 0.6

```{r}
results[7,] = data.frame(getTrainPerf(wave_0.6), Name = paste0("wave_0.6"), Details = "WaveICA2.0 Cutoff 0.6")

error1<-wave_0.6$pred %>%
  filter(alpha == wave_0.6$bestTune$alpha & lambda == wave_0.6$bestTune$lambda) %>% #subset 
  separate(Resample, c("fold", "rep"), "\\.") %>% 
  group_by(rep) %>% #group by replicate
  summarise(RMSE = RMSE(obs, pred)) %>% 
  as.data.frame() %>% 
  mutate(Model = "wave_0.6", Details = "WaveICA2.0 Cutoff 0.6")

error<-rbind(error,error1)
```

## WaveICA2.0 Cutoff of 0.7

```{r}
results[8,] = data.frame(getTrainPerf(wave_0.7), Name = paste0("wave_0.7"), Details = "WaveICA2.0 Cutoff 0.7")

error1<-wave_0.7$pred %>%
  filter(alpha == wave_0.7$bestTune$alpha & lambda == wave_0.7$bestTune$lambda) %>% #subset 
  separate(Resample, c("fold", "rep"), "\\.") %>% 
  group_by(rep) %>% #group by replicate
  summarise(RMSE = RMSE(obs, pred)) %>% 
  as.data.frame() %>% 
  mutate(Model = "wave_0.7", Details = "WaveICA2.0 Cutoff 0.7")

error<-rbind(error,error1)
```

## WaveICA2.0 Cutoff of 0.8

```{r}
results[9,] = data.frame(getTrainPerf(wave_0.8), Name = paste0("wave_0.8"), Details = "WaveICA2.0 Cutoff 0.8")

error1<-wave_0.8$pred %>%
  filter(alpha == wave_0.8$bestTune$alpha & lambda == wave_0.8$bestTune$lambda) %>% #subset 
  separate(Resample, c("fold", "rep"), "\\.") %>% 
  group_by(rep) %>% #group by replicate
  summarise(RMSE = RMSE(obs, pred)) %>% 
  as.data.frame() %>% 
  mutate(Model = "wave_0.8", Details = "WaveICA2.0 Cutoff 0.8")

error<-rbind(error,error1)
```

## WaveICA2.0 Cutoff of 0.9

```{r}
results[10,] = data.frame(getTrainPerf(wave_0.9), Name = paste0("wave_0.9"), Details = "WaveICA2.0 Cutoff 0.9")

error1<-wave_0.9$pred %>%
  filter(alpha == wave_0.9$bestTune$alpha & lambda == wave_0.9$bestTune$lambda) %>% #subset 
  separate(Resample, c("fold", "rep"), "\\.") %>% 
  group_by(rep) %>% #group by replicate
  summarise(RMSE = RMSE(obs, pred)) %>% 
  as.data.frame() %>% 
  mutate(Model = "wave_0.9", Details = "WaveICA2.0 Cutoff 0.9")

error<-rbind(error,error1)
```

## WaveICA2.0 Cutoff of 1

```{r}
results[11,] = data.frame(getTrainPerf(wave_1), Name = paste0("wave_1"), Details = "WaveICA2.0 Cutoff 1")

error1<-wave_1$pred %>%
  filter(alpha == wave_1$bestTune$alpha & lambda == wave_1$bestTune$lambda) %>% #subset 
  separate(Resample, c("fold", "rep"), "\\.") %>% 
  group_by(rep) %>% #group by replicate
  summarise(RMSE = RMSE(obs, pred)) %>% 
  as.data.frame() %>% 
  mutate(Model = "wave_1", Details = "WaveICA2.0 Cutoff 1")

error<-rbind(error,error1)
```

## Making the cool plots
```{r fig.asp = 0.9, fig.width = 7}
error.p<-error %>% ggplot(aes(x=fct_inorder(Details),y=RMSE, color=Model))+
  geom_boxplot(outlier.colour = "red", alpha = 0.3)+
  geom_boxplot(outlier.colour="red", outlier.shape=1,
                outlier.size=8)+
  geom_jitter(shape=16, position=position_jitter(0))+
  theme(legend.position="top")+
  theme(axis.title.x = element_blank())

table.p <- ggtexttable(results, rows = NULL, 
                        theme = ttheme("mOrange", base_size = 7))
ggarrange(error.p, table.p, 
          ncol = 1, nrow = 2,
          heights = c(1, 0.5, 0.3))
```

```{r fig.asp = 0.9, fig.width = 8}
error.p<-error %>% slice(26:nrow(error)) %>% ggplot(aes(x=fct_inorder(Details),y=RMSE, color=Model))+
  geom_boxplot(outlier.colour = "red", alpha = 0.3)+
  geom_boxplot(outlier.colour="red", outlier.shape=1,
                outlier.size=8)+
  geom_jitter(shape=16, position=position_jitter(0))+
  theme(legend.position="top")+
  theme(axis.title.x = element_blank())

resultstemp<-results %>% slice(2:nrow(results))
table.p <- ggtexttable(resultstemp, rows = NULL, 
                        theme = ttheme("mOrange", base_size = 7))
rm(resultstemp)
ggarrange(error.p, table.p, 
          ncol = 1, nrow = 2,
          heights = c(1, 0.5, 0.3))
```
